---
title: "Class 05: Data Visualization with GGPLOT"
author: "sylvia ho"
format: pdf
---

## background ##
there are lots of ways to make plots in R these include so called base R like the `plot()` and add packages like **ggplot2**

same plot diff sysyems use `cars`
```{r}
head(cars)
```
```{r}
plot(cars)
```

first instiall `install.packages("ggplot2")`
>**never run an install packages in code chunk bc will re install every render**
when use add-on pkg need to load w `library()`
every ggplot needs 3 things
- data (eg data frame)
- aes (map data to plot)
- geom (plot type, lines etc)

```{r}
library(ggplot2)

#coment 

ggplot(cars) + 
  aes(x=speed, y=dist) +
  geom_point() +
  labs(title="Speed and Stopping Distances of Cars",
       x="Speed (MPH)", 
       y="Stopping Distance (ft)",
       subtitle = "ft vs mph",
       caption="Dataset: 'cars'") +
  geom_smooth(method="lm", se=FALSE) +
  theme_bw()
```

aes
```{r}
url <- "https://bioboot.github.io/bimm143_S20/class-material/up_down_expression.txt"
genes <- read.delim(url)
head(genes)
nrow(genes)
colnames (genes)
ncol(genes)
table(genes$State)
127/5196

```
plot

```{r}
p <- ggplot(genes) + 
    aes(x=Condition1, y=Condition2, col=State) +
    geom_point()
p + scale_colour_manual(values=c("skyblue3","gray","pink1")) +
    labs(title="Gene Expression Changes Upon Drug Treatment",
         x="Control (no drug) ",
         y="Drug Treatment")


```
gapminder install
`install.packages("gapminder")`

```{r}
library(gapminder)
```
- dplyr *itals* **bold**

```{r}
# install.packages("dplyr")  ## un-comment to install if needed
library(dplyr)

gapminder_2007 <- gapminder %>% filter(year==2007)

ggplot(gapminder_2007) +
  geom_point(aes(x=gdpPercap, y=lifeExp, color = continent, size = pop), alpha=.5) +
scale_size_area(max_size = 10)

```

1957
```{r}

gapminder_1957 <- gapminder %>% filter(year==1957 | year==2007)

ggplot(gapminder_1957) +
  geom_point(aes(x=gdpPercap, y=lifeExp, color = continent, size = pop), alpha=.7) +
scale_size_area(max_size = 15)+
  facet_wrap(~year)
```

bars
```{r}
gapminder_top5 <- gapminder %>% 
  filter(year==2007) %>% 
  arrange(desc(pop)) %>% 
  top_n(5, pop)

gapminder_top5

ggplot(gapminder_top5) +
  aes(x=reorder(country, -pop), y=pop, fill=gdpPercap) +
  geom_col()

```

```{r}
USArrests$State <- rownames(USArrests)
ggplot(USArrests) +
  aes(x=reorder(State,Murder), y=Murder) +
  geom_col() +
  coord_flip()

ggplot(USArrests) +
  aes(x=reorder(State,Murder), y=Murder) +
  geom_point() +
  geom_segment(aes(x=State, 
                   xend=State, 
                   y=0, 
                   yend=Murder), color="blue") +
  coord_flip()
```
ani
`install.packages("gifski") 
install.packages("gganimate")`
```{r}

library(gapminder)
library(gganimate)
"
# Setup nice regular ggplot of the gapminder data
ggplot(gapminder, aes(gdpPercap, lifeExp, size = pop, colour = country)) +
  geom_point(alpha = 0.7, show.legend = FALSE) +
  scale_colour_manual(values = country_colors) +
  scale_size(range = c(2, 12)) +
  scale_x_log10() +
  # Facet by continent
  facet_wrap(~continent) +
  # Here comes the gganimate specific bits
  labs(title = 'Year: {frame_time}', x = 'GDP per capita', y = 'life expectancy') +
  transition_time(year) +
  shadow_wake(wake_length = 0.1, alpha = FALSE)
  "
```

patchwork
`install.packages("patchwork") `
```{r}

library(patchwork)

# Setup some example plots 
p1 <- ggplot(mtcars) + geom_point(aes(mpg, disp))
p2 <- ggplot(mtcars) + geom_boxplot(aes(gear, disp, group = gear))
p3 <- ggplot(mtcars) + geom_smooth(aes(disp, qsec))
p4 <- ggplot(mtcars) + geom_bar(aes(carb))

# Use patchwork to combine them here:
(p1 | p2 ) /
     (p3|p4)
```

```{r}
sessionInfo()

```

