---
title: "class06: R functions"
author: "Sylvia Ho (PID: a18482382)"
format: pdf
---

## background
functions are at the heart of using r everything insvolves calling and using functions
all fns have at least 3 thigns:

1. a **name** the thing we use to call it
2. input **args** that are comma separated
3. **body** lines of code btwn curly {} that does work of fn

```{r}
add<- function(x){
  x+1
}
```

```{r}
add(c(100,200,300))
```
```{r}
add<- function(x,y){
  x+y
}
```


```{r}

add(100,10)
```
```{r}

log(10,base=10)
```

>**n.b** input args can be either reqd or opt the latter have fall back default that is spec in fn w an equal sign

```{r}
#add(x=100,y=200,z=300)
```

## a second fn

all fns in R look like this

```name<fn(arg){
body
}
```

the`sample()` fn in R samples randomly from arg
```{r}
sample(1:10)

```
>Q writea first version fn called `generate dna()` that gens a user spec length n random dna seq

```{r}

generate_dna<-function(n=6){
  bases<-c("a","t","g","c")
  sample(bases, size=n, replace = TRUE)
}

generate_dna(100)

```
>Q mod to return actual seq

```{r}

generate_dna<-function(n=6){
  bases<-c("a","t","g","c")
  ans<-sample(bases, size=n, replace = TRUE)
 an<-paste(ans,collapse="")
 return(an)
}

generate_dna(10)
```

>Q option FASTA or vector

```{r}

generate_dna<-function(n=6, fasta = TRUE){
  bases<-c("a","t","g","c")
  ans<-sample(bases, size=n, replace = TRUE)
 
 if (fasta){
   ans<-paste(ans,collapse="")

 }
 return(ans)
 
}
```

```{r}
generate_dna(10)
generate_dna(10, fasta = F)
```

```{r}
generate_protein<- function(){
  aa<-c("A", "R", "N", "D", "C", "E", "Q", "G", "H", "I",
"L", "K", "M", "F", "P", "S", "T", "W", "Y", "V")
  
}

```

```{r}

generate_protein<-function(n=6, fasta = TRUE){
aa<-c("A", "R", "N", "D", "C", "E", "Q", "G", "H", "I",
"L", "K", "M", "F", "P", "S", "T", "W", "Y", "V")
  
  ans<-sample(aa, size=n, replace = TRUE)
 
 if (fasta){
   ans<-paste(ans,collapse="")

 }
 return(ans)
 
}
generate_protein(10, fasta = F)
```

>Q gen protein gn to gen seqs between 6 to 12, check in uniqueue aka against NR database at NCBI

```{r}

generate_protein<-function(n=6, fasta = TRUE){
aa<-c("A", "R", "N", "D", "C", "E", "Q", "G", "H", "I",
"L", "K", "M", "F", "P", "S", "T", "W", "Y", "V")
  
  ans<-sample(aa, size=n, replace = TRUE)
 
 if (fasta){
   ans<-paste(ans,collapse="")

 }
 return(ans)}
```
   
  
  
```{r}


for (i in 6:12) {
  cat(">",i, "\n")
  cat(generate_protein(i), "\n")
}
```
